---
title: "Korea Social Contacts"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages

```{r}
library(data.table)
library(dplyr)
library(sf)
library(DT)
library(ggplot2)
```

Function to label region names in English

```{r, warning=FALSE, message=FALSE}
kor <- st_read("data/gadm41_KOR_1.shp")
# kor$NAME_1
# 1	서울특별시
# 2	부산광역시
# 3	대구광역시
# 4	인천광역시
# 5	광주광역시
# 6	대전광역시
# 7	울산광역시
# 8	세종특별자치시
# 9	경기도
# 10	강원도
# 11	충청북도
# 12	충청남도
# 13	전라북도
# 14	전라남도
# 15	경상북도
# 16	경상남도
# 17	제주특별자치도
kor_regions <- c("Seoul",
                 "Busan","Daegu","Incheon","Gwangju","Daejeon","Ulsan",
                 "Sejong",
                 "Gyeonggi-do","Gangwon-do",
                 "Chungcheongbuk-do","Chungcheongnam-do",
                 "Jeollabuk-do","Jeollanam-do",
                 "Gyeongsangbuk-do","Gyeongsangnam-do",
                 "Jeju")

get_region <- function(i, regions){
  return(regions[i])
}
```

### Data 

#### Respondents

##### Check #1: File for the survey results
```{r}
res <- fread("data/df_res_cleansing_ver_3.csv")
unique(res$SQ3)
res$age <- as.integer(res$SQ1)
res$gender <- ifelse(res$SQ2 == 1, "M", "F")
res$region <- get_region(res$SQ3, region=kor_regions)
res$hhsize <- as.integer(res$DQ2) # household size
res$id <- res$ID_ori
```

#### Contacts

##### Check #2: File for the contacts
```{r}
cont <- fread("data/df_con_cleansing_ver_3.csv")
# datatable(cont) # interactive table exploration
unique(cont$접촉날짜)
cont$date <- ifelse(cont$접촉날짜 %in% c("1206","1207","1208","1209","1210","1211","1212"),
                   paste0("2023",cont$접촉날짜), paste0("20240",cont$접촉날짜))
cont$date <- as.Date(cont$date, format="%Y%m%d")
cont$dayofweek <- weekdays(cont$date)
cont$dayofweek <- factor(cont$dayofweek, levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
cont$id <- cont$ID_ori
cont$duration <- as.integer(cont$Q8) # timehousehold size

cont_1st <- cont[cont$date < as.Date("2024-01-01")] # first survey
cont_2nd <- setdiff(cont, cont_1st) #2nd survey

sprintf("Total respondents: %d (1st) and %d (2nd)",
        length(unique(cont_1st$ID)), length(unique(cont_2nd$ID)))
```

##### Check #3: Number of contacts
```{r}
cont_1st |> 
  dplyr::group_by(id, date, dayofweek) |> 
  dplyr::summarize(contacts = max(접촉순서),
                   mean_duration = mean(duration)) -> cont_1st_n 
```

Combine the participants and contacts data

```{r}
dat <- left_join(res, cont_1st_n, by="id")
```

Age group classification

##### Check #4: Age classification: 0-9, 10-19, ... 
```{r}
age_grp_label <- c("0-9","10-19","20-29","30-39","40-49","50-59","60-69","70+")
sum(is.na(dat$age))
dat$age_grp <- ifelse(dat$age < 70, age_grp_label[floor(dat$age/10) + 1], 
                      age_grp_label[8])
dat$age_grp <- factor(dat$age_grp, levels=c(age_grp_label))
table(dat$age_grp)
```

##### Check #4: Age classification: 0-4, 5-9, 10-14, 15-19, 20-29, 30-39, ..., 70+ 
```{r}
age_grp_label <- c("0-4","5-9","10-14","15-19","20-29",
                   "30-39","40-49","50-59","60-69","70+")

dat$age_grp <- ifelse(is.na(dat$age), "Missing", 
                      ifelse(dat$age < 20,
                             age_grp_label[floor(dat$age/5) + 1], 
                             ifelse(dat$age < 80,
                                    age_grp_label[floor(dat$age/10) + 3],
                                    age_grp_label[10])))
dat$age_grp <- factor(dat$age_grp, levels=c(age_grp_label))
table(dat$age_grp)
```

Household size

##### Check #5: Household size classification: 1, 2, .., 5+ 
```{r}
unique(dat$hhsize)
table(dat$hhsize)
sum(is.na(dat$hhsize))
dat$hhsize_grp <- ifelse(dat$hhsize >= 5, "5+", as.character(dat$hhsize))
dat$hhsize_grp <- factor(dat$hhsize_grp, levels=c(as.character(paste0(1:4)),"5+"))
table(dat$hhsize_grp)
```

### Descriptive statistics 


Contacts

```{r}
summary(dat$contacts)
table(dat$contacts)
```

Number of contacts by the day of the week

```{r}
library(ggplot2)
dat |> 
  ggplot()+
  geom_histogram(aes(x=contacts))+
  facet_wrap(~dayofweek)+
  theme_bw()

dat |> 
  ggplot()+
  geom_violin(aes(x=dayofweek, y=contacts))

dat |> 
  dplyr::group_by(dayofweek) |> 
  dplyr::summarize(mean_contacts = mean(contacts),
                   se_contacts = sd(contacts)/sqrt(n()),
                   lb_contacts = mean_contacts - 1.96 * se_contacts,
                   ub_contacts = mean_contacts + 1.96 * se_contacts)
```

Number of contacts and contact duration by age group

```{r}
table(dat$age_grp)

dat |> 
  ggplot()+
  geom_histogram(aes(x=contacts))+
  facet_wrap(~age_grp)+
  theme_bw()

dat |> 
  ggplot()+
  geom_violin(aes(x=age_grp, y=contacts))

dat |> 
  dplyr::group_by(age_grp) |> 
  dplyr::summarize(mean_contacts = mean(contacts),
                   se_contacts = sd(contacts)/sqrt(n()),
                   lb_contacts = mean_contacts - 1.96 * se_contacts,
                   ub_contacts = mean_contacts + 1.96 * se_contacts)

# duration is a category though.
dat |> 
  dplyr::group_by(age_grp) |> 
  dplyr::summarize(mean_duration_ = mean(mean_duration),
                   se_duration = sd(mean_duration)/sqrt(n()),
                   lb_duration = mean_duration_ - 1.96 * se_duration,
                   ub_duration = mean_duration_ + 1.96 * se_duration)

```

Mean number of contact by household size

```{r}
table(dat$hhsize_grp)

dat |> 
  dplyr::group_by(hhsize_grp) |> 
  dplyr::summarize(mean_contacts = mean(contacts),
                   se_contacts = sd(contacts)/sqrt(n()),
                   lb_contacts = mean_contacts - 1.96 * se_contacts,
                   ub_contacts = mean_contacts + 1.96 * se_contacts)
```

Mean number of contact by region

```{r}
table(dat$hhsize_grp)

dat |> 
  dplyr::group_by(region) |> 
  dplyr::summarize(mean_contacts = mean(contacts),
                   se_contacts = sd(contacts)/sqrt(n()),
                   lb_contacts = mean_contacts - 1.96 * se_contacts,
                   ub_contacts = mean_contacts + 1.96 * se_contacts)
```


Bivariate relationship

```{r}
library(Hmisc)
vars_of_interest <- c("contacts", "age_grp", "gender", "dayofweek", "hhsize_grp")
d <- dat[,..vars_of_interest]
GGally::ggpairs(d)
```

Correlation
```{r}
# glimpse(dat)
# voi <- c("age", "hhsize", "contacts")
# Hmisc::rcorr(as.matrix(dat[,..voi]))
# cor.test(dat[,..voi])
library("ggpubr")
ggscatter(dat, x = "age", y = "contacts", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Age", ylab = "Number of contacts")

ggscatter(dat, x = "hhsize", y = "contacts", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Household size", ylab = "Number of contacts")

# ggscatter(dat, x = "dayofweek", y = "contacts", 
#           add = "reg.line", conf.int = TRUE, 
#           cor.coef = TRUE, cor.method = "pearson",
#           xlab = "Day of week", ylab = "Number of contacts")
```


### Regresssion analysis

Label variables as factors
```{r}
dat$dayofweek <- factor(dat$dayofweek, levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
dat$age_grp <- factor(dat$age_grp, levels=c(age_grp_label))
dat$hhsize_grp <- factor(dat$hhsize_grp, levels=c(as.character(paste0(1:4)),"5+"))
dat$hhsize_grp <- relevel(dat$hhsize_grp, ref="1")
dat$gender <- factor(dat$gender, levels=c("F","M"))
dat$region <- factor(dat$region, levels=kor_regions)
```

Check completeness of the data

```{r}
# model_vars <- c("contacts", "age_grp", "gender", "dayofweek", 
               # "hhsize_grp", "region")
# dat <- dat[,..model_vars]
all.equal(dat, dat[!is.na(contacts),]) # no NA's
```

### Regression

Negative Binomial Regression 

```{r}
library(MASS)
summary(m1 <- glm.nb(contacts ~ age_grp + gender + dayofweek + hhsize_grp                      + region, data = dat))

knitr::kable(exp(m1$coefficients))
```

Poisson Regression

```{r}
library(MASS)
summary(m2 <- glm(contacts ~ age_grp + gender + dayofweek + hhsize_grp + region,
                  family="poisson", data = dat))
# exp(m2$coefficients)
df <- data.frame(Parameters=names(coef(m1)), NegBin=exp(as.numeric(coef(m1))),
                 Poisson=exp(as.numeric(coef(m2))))
knitr::kable(df)
```

Model comparison
```{r}
AIC(m1,m2)
```


```{r}
summary(dat$contacts)
table(dat$contacts)
```

### Weighting

```{r}
dat$age_grp_wt_row <- as.integer(ifelse(dat$age < 70, (floor(dat$age/10) + 1)*3, 8*3))
table(dat$age_grp)
table(dat$age_grp_wt_row)

dat$hhsize_grp_wt_col <- as.integer(ifelse(dat$hhsize >= 5, 6, dat$hhsize + 1))
table(dat$hhsize_grp)
table(dat$hhsize_grp_wt_col)

wt_data <- as.data.frame(fread("data/ratio_5plus.csv"))
dat$wt <- sapply(1:nrow(dat), 
                 function(i) as.numeric(wt_data[dat$age_grp_wt_row[i], dat$hhsize_grp_wt_col[i]]))

```

Negative Binomial Regression with weight  

```{r}
summary(m3 <- glm.nb(contacts ~ age_grp + gender + dayofweek + hhsize_grp + region,
                     data = dat, weights = wt))
# exp(m1$coefficients)
df <- data.frame(Parameters=names(coef(m1)), NegBin=exp(as.numeric(coef(m1))),
                 NegBin_Weight=exp(as.numeric(coef(m3))))
knitr::kable(df)
```

```{r}
summary(m4 <- glm.nb(contacts ~ age_grp + gender + dayofweek + hhsize_grp,
                     data = dat, weights = wt))
# exp(m1$coefficients)
# df <- data.frame(Parameters=names(coef(m1)), NegBin=exp(as.numeric(coef(m1))),
#                  NegBin_Weight=exp(as.numeric(coef(m3))))
# knitr::kable(df)
```

Create table
```{r}
## Estimates
xhat <- exp(summary(m4)$coefficients[,1])
lb <- exp(summary(m4)$coefficients[,1]) - 
  qnorm(0.975) * summary(m4)$coefficients[,2] 
ub <- exp(summary(m4)$coefficients[,1]) + 
  qnorm(0.975) * summary(m4)$coefficients[,2]

est_ <- sprintf("%.2f (%.2f - %.2f)", xhat, lb, ub)
# include reference values
est <- c(est_[1], "1.00", est_[2:10], "1.00", est_[11], "1.00", est_[12:17],
"1.00",est_[18:21])

## Number of participants
dat |> 
  dplyr::group_by(age_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> age_counts

dat |> 
  dplyr::group_by(gender) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> sex_counts

dat |> 
  dplyr::group_by(dayofweek) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> dayofweek_counts

dat |> 
  dplyr::group_by(hhsize_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> household_counts

numpart <- c(age_counts$n, sex_counts$n, dayofweek_counts$n, household_counts$n)
mean_contacts <- c(age_counts$mean_contacts, sex_counts$mean_contacts, dayofweek_counts$mean_contacts, household_counts$mean_contacts)
sd_contacts <- c(age_counts$sd_contacts, sex_counts$sd_contacts, dayofweek_counts$sd_contacts, household_counts$sd_contacts)
contacts <- sprintf("%.2f (%.2f)", mean_contacts, sd_contacts)

covars <- c("0-4", "5-9", "10-14", "15-19", "20-29","30-39", "40-49",
               "50-59", "60-69", "70+",
               "Female", "Male",
               "Sunday","Monday","Tuesday","Wednesday","Thursday","Friday",
               "Saturday", 
               "1", "2", "3", "4", "5+")

df <- data.frame(category=c(rep("Age of participants",10), 
                          rep("Sex of participant",2),
                          rep("Day of the week",7), rep("Household size",5)),
                 covariate = covars, 
                 num = numpart,
                 marginals = contacts,
                 estimates = est[2:25])

names(df) <- c("Category","Covariate", "Number of participants", 
               "Marginal mean reported contacts (standard deviation)",
               "Estimates")

knitr::kable(df)
m4$theta
m4$SE.theta
est <- m3$theta
lb <- m4$theta - qnorm(0.975)*m4$SE.theta 
ub <- m4$theta + qnorm(0.975)*m4$SE.theta 
sprintf("Dispersion parameter, theta = %.2f (%.2f - %.2f)", est, lb, ub)
sprintf("alpha = %.2f (%.2f - %.2f)", 1/est, 1/ub, 1/lb)
```

In the negative binomial regression, the variance is given as $\mu+\frac{\mu^2}{\theta}$. Most other software, however, use the direct parameterization $\mu+\alpha\mu^2$. So did the POLYMOD paper, in which  $\alpha=0.36~(95\text{%}~\text{CI}:0.34-0.37)$
 
Household size distribution from the census
```{r, eval=F, include=F}
library(readxl)
hhsize <- read_xlsx("data/2023년 6월 주민등록_세대원수별_가구수.xlsx")
nms <- hhsize[2,]
hhsize <- hhsize[3:nrow(hhsize),]
names(hhsize) <- as.character(nms)
for (i in 3:ncol(hhsize)) {
  hhsize[,i] <- as.integer(hhsize[,i,drop=TRUE])
}
hhsize$region <- c("Korea", kor_regions)
korea_hhsize_prop <- unlist(hhsize[1, 4:13]) / unlist(hhsize[1, 3])
sum(korea_hhsize_prop)

dat |> 
  dplyr::group_by(hhsize_grp) |> 
  dplyr::summarize(num_hh = n()) -> dat_hhsize

dat_hhsize$prop <- dat_hhsize$num_hh / sum(dat_hhsize$num_hh)

df <- data.frame(hhsize_grp = c(as.character(paste0(1:9)),"10+"), 
                 prop_census = korea_hhsize_prop)
dat_hhsize <- left_join(df, dat_hhsize, by="hhsize_grp")
dat_hhsize$num_hh <- ifelse(is.na(dat_hhsize$num_hh), 0, dat_hhsize$num_hh)
dat_hhsize$prop <- ifelse(is.na(dat_hhsize$prop), 0, dat_hhsize$prop)

dat_hhsize$wt <- dat_hhsize$prop_census / dat_hhsize$prop
dat_hhsize
```

Age group distribution from the census
```{r, eval=FALSE, include=FALSE}
agedist <- read_xlsx("data/2023년_6월_주민등록_거주자_인구_ver_2.xlsx")
ad <- agedist[4:nrow(agedist), 2:12]
pop <- sapply(2:11, function(x) as.numeric(gsub(",", "", ad[,x,drop=T])))
ad2 <- cbind(ad[,1], pop)
names(ad2) <- names(ad)
ad3 <- ad2[,2:11]/ad2[,3] # proportion of each age group
ad3$sum <- rowSums(ad3[,3:10])

ad4 <- cbind(data.frame(region=c("Korea", kor_regions), ad3))
names(ad4) <- c("Region", "residents", "Census_pop", age_grp_label, "Sum_prop")

# sample age dist
dat |> 
  dplyr::group_by(age_grp) |> 
  dplyr::summarize(pop = n()) -> dat_agedist

dat_agedist$prop <- dat_agedist$pop / sum(dat_agedist$pop)
dat_agedist$prop_census <- unlist(ad4[1, age_grp_label])
dat_agedist$wt <-  dat_agedist$prop_census / dat_agedist$prop
dat_agedist
```

