---
title: "Korea Social Contacts: Linear Regression"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: html
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Packages
```{r, echo=F}
library(data.table)
library(MASS)
library(lme4)
library(tidyverse)
source("utils.R")
```

## Data
```{r}
dat <- data.table::fread("outputs/contact_data_20240528.csv")
dat$date <- as.Date(dat$date)
dat$date_grp <- factor(dat$date_grp, levels=c("Weekdays", "WeekdaysVacation","Weekends","LunarNewYear"))

dat$dayofweek <- factor(dat$dayofweek, 
                        levels=c("Sunday","Monday","Tuesday","Wednesday",
                                 "Thursday","Friday","Saturday"))
age_grp_label <- c("0-4","5-9","10-14","15-19","20-29",
                   "30-39","40-49","50-59","60-69","70-79")
dat$age_grp <- factor(dat$age_grp, levels=age_grp_label)

age_grp_broader_label <- c("0-4","5-9","10-19","20-29",
                   "30-39","40-59","60-79")
dat$age_grp_broader <- factor(dat$age_grp_broader, levels=age_grp_broader_label)
dat$hhsize_grp <- factor(dat$hhsize_grp, 
                         levels=c(as.character(paste0(1:4)),"5+"))
dat$hhsize_grp <- relevel(dat$hhsize_grp, ref="1")
dat$sex <- factor(dat$sex, levels=c("F","M"))
kor_regions <- c("Seoul",
                 "Busan","Daegu","Incheon","Gwangju","Daejeon","Ulsan",
                 "Sejong",
                 "Gyeonggi-do","Gangwon-do",
                 "Chungcheongbuk-do","Chungcheongnam-do",
                 "Jeollabuk-do","Jeollanam-do",
                 "Gyeongsangbuk-do","Gyeongsangnam-do",
                 "Jeju")
dat$region <- factor(dat$region)

# dat_full <- dat
# dat <- subset(dat_full, date < as.Date("2024-01-01")) # first survey
```

## Summary statistics
Some crude statistics to compare with the regression outputs
```{r}
dat |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   age_grp = age_grp[1]) |> 
  dplyr::group_by(age_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> age_counts

dat |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   sex = sex[1]) |> 
  dplyr::group_by(sex) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> sex_counts


dat |> 
  dplyr::group_by(date_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> date_grp_counts

ndays_dategrp <- c(5,3,2,4) # number of days for each date group to adjust the number of individuals
date_grp_counts$n <-  date_grp_counts$n / ndays_dategrp

dat |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   hhsize_grp = hhsize_grp[1]) |> 
  dplyr::group_by(hhsize_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> household_counts
```

```{r, eval=F}
library(brms)
fit <- readRDS("outputs/baye_dategrp_20240607.rds")
summary(fit)
ranef(fit)
fixef(fit)[,1]
```

### POLYMOD-like table
```{r}
## Estimates
fit <- readRDS("outputs/baye_dategrp_20240607.rds")
# fit <- readRDS("outputs/baye_dategrp_pois_20240609.rds")
xhat <- exp(fixef(fit)[,1])
lb <- exp(fixef(fit)[,3])
ub <- exp(fixef(fit)[,4])

est_ <- sprintf("%.2f (%.2f - %.2f)", xhat, lb, ub)
# include reference values
est <- c(est_[1], "1.00", est_[2:10], "1.00", est_[11], "1.00", est_[12:14],
"1.00",est_[15:18])


numpart <- c(age_counts$n, sex_counts$n, date_grp_counts$n, 
             household_counts$n)
mean_contacts <- c(age_counts$mean_contacts, sex_counts$mean_contacts,
                   date_grp_counts$mean_contacts, 
                   household_counts$mean_contacts)
sd_contacts <- c(age_counts$sd_contacts, sex_counts$sd_contacts,
                 date_grp_counts$sd_contacts, household_counts$sd_contacts)
crude_ratio <- c(mean_contacts[1:10]/mean_contacts[1],
                 mean_contacts[11:12]/mean_contacts[11],
                 mean_contacts[13:16]/mean_contacts[13],
                 mean_contacts[17:21]/mean_contacts[17])
crude_ratio <- formatC(crude_ratio, digits = 2)

contacts <- sprintf("%.2f (%.2f)", mean_contacts, sd_contacts)

covars <- c("0-4", "5-9", "10-14", "15-19", "20-29","30-39", "40-49",
               "50-59", "60-69", "70+",
               "Female", "Male",
               "Weekdays", "Vacation weekdays", "Weekends", "New Year Holidays",
               "1", "2", "3", "4", "5+")

df <- data.frame(category=c(rep("Age",10), 
                          rep("Sex",2),
                          rep("Date group",4), rep("HH size",5)),
                 covariate = covars, 
                 num = numpart,
                 marginals = contacts,
                 crude_ratio = crude_ratio,
                 estimates = est[2:22]) # exclude intercept

names(df) <- c("Cat", "Covar", "No. part", 
               "Mar mean (sd)","CR",
               "Est. (95% CrI)")

knitr::kable(df)
knitr::kable(df, format = "markdown")
# random effect
rand_eff <- sprintf("%.2f (%.2f - %.2f)", 0.48, 0.46, 0.50)
phi <- sprintf("Dispersion parameter, phi = %.2f (%.2f - %.2f)", 14.03, 13.19, 14.94)

# sprintf("Dispersion parameter, phi = %.2f (%.2f - %.2f)", est, lb, ub)
# sprintf("alpha = %.2f (%.2f - %.2f)", 1/est, 1/ub, 1/lb)
```

Intraclass correlation (ICC)
```{r eval=FALSE}
variance_decomposition(fit)
# # Random Effect Variances and ICC
# 
# Conditioned on: all random effects
# 
# ## Variance Ratio (comparable to ICC)
# Ratio: 0.46  CI 95%: [0.42 0.50]
# 
# ## Variances of Posterior Predicted Distribution
# Conditioned on fixed effects:  8.60  CI 95%: [ 8.00  9.29]
# Conditioned on rand. effects: 16.04  CI 95%: [15.49 16.61]
# 
# ## Difference in Variances
# Difference: 7.44  CI 95%: [6.62 8.22]> variance_decomposition(fit)
# # Random Effect Variances and ICC
# 
# Conditioned on: all random effects
# 
# ## Variance Ratio (comparable to ICC)
# Ratio: 0.46  CI 95%: [0.42 0.50]
# 
# ## Variances of Posterior Predicted Distribution
# Conditioned on fixed effects:  8.60  CI 95%: [ 8.00  9.29]
# Conditioned on rand. effects: 16.04  CI 95%: [15.49 16.61]
# 
# ## Difference in Variances
# Difference: 7.44  CI 95%: [6.62 8.22]
```


### Posterior predictive check
```{r}
library(brms)
pp_check(fit, type = "bars", ndraws = ) + xlim(0, max(dat$contacts))
pp_check(m, type = "bars_grouped", group = "group") + xlim(min, max)

pp <- posterior_predict(fit, ndraws=1000)
# data.table::fwrite(pp, "outputs/nb_mix_dategrp_post_pred.csv")
dim(pp)

get_freq <- function(d) {
  data.frame(table(factor(d, levels = c(0:50))))[,2]
}
df <- data.frame(matrix(NA, nrow=51, ncol=1001))

for (i in 1:1000) {
  df[, i+1] <- get_freq(pp[i,])
}
df <- df[,2:51]
dfs <- apply(df, 1, quantile, probs=c(0.025,0.5,0.975))
dfss <- data.frame(x=0:50, data.frame(t(dfs)))             
names(dfss) <- c("x","2.5%","50%","97.5%")
library(ggplot2)
ggplot() +
  geom_histogram(data=dat, aes(x=contacts), bins=51, fill="grey40") +
  geom_point(data=dfss, aes(x=x, y=`50%`), color="firebrick",fill="firebrick",size=1.5) +
  geom_errorbar(data=dfss, aes(x=x, ymin=`2.5%`, ymax=`97.5%`), width=0, 
                color="firebrick", linewidth=0.4) +
  labs(X="Contacts", y="Frequency")+
  theme_bw() -> plt
fac <- 1.5
ggsave("plots/pp_check_pois.png", plot=plt, width=3.4*fac, height=2.7*fac, units="in")  
```


$$
\text{NegBinomial}2(u|\mu,\phi)=\left(\begin{array}
 + y + \phi - 1 \\ 
 y 
 \end{array} \right) \left(\frac{\mu}{\mu+\phi} \right)^y \left(\frac{\phi}{\mu+\phi} \right)^\phi
$$
$$\bf{E}[Y]= \mu \text{ and }\bf{Var}[Y] = \mu + \frac{\mu^2}{\phi}
$$

Some crude statistics to compare with the regression outputs
```{r}
dat2023 <- subset(dat, date < as.Date("2024-01-01")) # first survey
dat2023 |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   age_grp = age_grp[1]) |> 
  dplyr::group_by(age_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> age_counts

dat2023 |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   sex = sex[1]) |> 
  dplyr::group_by(sex) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> sex_counts


dat2023 |> 
  dplyr::group_by(dayofweek) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts),
                   sd_contacts = sd(contacts))  -> dayofweek_counts

ndays_week <- 7 # number of days for each date group to adjust the number of individuals
date_grp_counts$n <-  date_grp_counts$n / ndays_week

dat2023 |> 
  dplyr::group_by(id) |>
  dplyr::summarize(contacts_id = mean(contacts),
                   hhsize_grp = hhsize_grp[1]) |> 
  dplyr::group_by(hhsize_grp) |>
  dplyr::summarize(n = n(),
                   mean_contacts = mean(contacts_id),
                   sd_contacts = sd(contacts_id))  -> household_counts
```


### POLYMOD-like table: Day of week 
```{r}
## Estimates

fit <- readRDS("outputs/baye_dayofweek_20240606.rds")
xhat <- exp(fixef(fit)[,1])
lb <- exp(fixef(fit)[,3])
ub <- exp(fixef(fit)[,4])

est_ <- sprintf("%.2f (%.2f - %.2f)", xhat, lb, ub)
# include reference values
est <- c(est_[1], "1.00", est_[2:10], "1.00", est_[11], "1.00", est_[12:17],
"1.00",est_[18:21])


numpart <- c(age_counts$n, sex_counts$n, dayofweek_counts$n, 
             household_counts$n)
mean_contacts <- c(age_counts$mean_contacts, sex_counts$mean_contacts,
                   dayofweek_counts$mean_contacts, 
                   household_counts$mean_contacts)
sd_contacts <- c(age_counts$sd_contacts, sex_counts$sd_contacts,
                 dayofweek_counts$sd_contacts, household_counts$sd_contacts)

crude_ratio <- c(mean_contacts[1:10]/mean_contacts[1],
                 mean_contacts[11:12]/mean_contacts[11],
                 mean_contacts[13:19]/mean_contacts[13],
                 mean_contacts[20:24]/mean_contacts[20])

crude_ratio <- formatC(crude_ratio, digits=2, format="f")
contacts <- sprintf("%.2f (%.2f)", mean_contacts, sd_contacts)

covars <- c("0-4", "5-9", "10-14", "15-19", "20-29","30-39", "40-49",
               "50-59", "60-69", "70+",
               "Female", "Male",
               "Sun","Mon","Tue","Wed","Thu","Fri",
               "Sat", 
               "1", "2", "3", "4", "5+")

df <- data.frame(category=c(rep("Age",10), 
                          rep("Sex",2),
                          rep("Day",7), rep("HH size",5)),
                 covariate = covars, 
                 num = numpart,
                 marginals = contacts,
                 crude_ratio = crude_ratio,
                 estimates = est[2:25]) # exclude intercept

names(df) <- c("Cat.","Covar", "No. part", 
               "Marginal mean (sd)",
               "CR",
               "Est.")

# knitr::kable(df)
knitr::kable(df, format = "markdown")
# random effect
rand_eff <- sprintf("%.2f (%.2f - %.2f)", 0.52, 0.50, 0.54)
phi <- sprintf("Dispersion parameter, phi = %.2f (%.2f - %.2f)", 
               52.03, 41.57, 66.47)

# m$theta
# m$SE.theta
# est <- m$theta
# lb <- m$theta - qnorm(0.975)*m$SE.theta 
# ub <- m$theta + qnorm(0.975)*m$SE.theta 
# sprintf("Dispersion parameter, theta = %.2f (%.2f - %.2f)", est, lb, ub)
# sprintf("alpha = %.2f (%.2f - %.2f)", 1/est, 1/ub, 1/lb)
```

### Regression
### Bayesian approach using the `brms` package 
```{r, eval=F}
library(brms)
# fit1 <- brms::brm(count ~ zAge + zBase * Trt + (1|patient),
#             data = epilepsy, family = poisson())
# summary(fit1)

# m_baye <- brms::brm(contacts ~ age_grp + date + hhsize_grp + (1|id),
#             data = dat, family = negbinomial())

m_baye <- brms::brm(contacts|weights(wt) ~ age_grp + dayofweek + hhsize_grp + (1|id), data = dat, family = negbinomial(),
            iter = 20000, warmup = 5000)

m_baye
# saveRDS(m_baye, paste0("outputs/baye_", tstamp(), ".rds"))
sm = summary(m_baye)
exp(sm$fixed[,"Estimate"])
# compare with m_nb_glmm
```

```{r}
m <- readRDS("outputs/baye_20240606.rds")
```

### Custom prior
```{r}
#
library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")
m0 <- brms::brm(contacts|weights(wt) ~ age_grp + sex + dayofweek + hhsize_grp + (1|id), data = dat, family = poisson(), 
                iter = 10000, 
                warmup = 5000, 
                cores = 4)
saveRDS(m0, paste0("outputs/baye_pois_dayofweek_", tstamp(), ".rds"))
Sys.time() - tic
```

Date group
```{r}
#
library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")
m0 <- brms::brm(contacts|weights(wt) ~ age_grp + sex + date_grp + hhsize_grp + (1|id), data = dat_full, family = negbinomial(), iter = 20000, warmup = 5000, cores = 4)
saveRDS(m0, paste0("outputs/baye_dategrp_", tstamp(), ".rds"))
Sys.time() - tic


library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")
m0 <- brms::brm(contacts|weights(wt) ~ age_grp + sex + date_grp + hhsize_grp + (1|id), data = dat_full, family = poisson(), iter = 20000, warmup = 5000, cores = 4)
saveRDS(m0, paste0("outputs/baye_dategrp_pois_", tstamp(), ".rds"))
Sys.time() - tic
```

```{r}
summary(m0)
ranef(m0)
pp_check(m0)
pp_check(m0, type = "error_hist", ndraws = 11)
pp_check(m0, type = "scatter_avg", ndraws = 100)
pp_check(m0, type = "stat_2d")
pp_check(m0, type = "rootogram")
pp_check(m0, type = "loo_pit")
```

### Fake-data simulation
```{r}
library(extraDistr)
N <- 1000
id <- 1:N
nrepeat <- 7
pop_intercept <- 1.3
# "0-4" "5-9" "10-14" "15-19" "20-29" "30-39" "40-49" "50-59" "60-69" "70-79"
age_grp <- sample(1:5, size=N, replace=TRUE) # with equal probability 
# c("Weekdays", "WeekdaysVacation", "Weekends", "LunarNewYear")
date_grp <- sample(1:4, size=N, replace=TRUE)
# c(as.character(1:4),"5+")
hhsize_grp <- sample(1:5, size=N, replace=TRUE, 
                     prob=c(0.2,0.25,0.25,0.25,0.05))

log_rr_age <- c(0, 0.41, 0.6, -0.42, -0.01)
log_rr_dategrp <- c(0, -0.41, -0.26, -0.15)
log_rr_hhsizegrp <- c(0, 0.21, 0.36, 0.45, 0.64)

random_error <- rnorm(N, mean=0, sd=0.02)

log_mean_contacts <- pop_intercept + log_rr_age[age_grp] + 
  log_rr_dategrp[date_grp] + log_rr_hhsizegrp[hhsize_grp] +
   + random_error

contacts <- rtpois(N, lambda=exp(log_mean_contacts), a=0, b=60)
wt <- rnorm(N, 1, 0.2)

d <- data.frame(id = as.character(id),
                contacts = contacts, 
                age_grp = as.factor(age_grp), 
                hhsize_grp = as.factor(hhsize_grp),
                date_grp = as.factor(date_grp),
                wt = wt)

library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_pois <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp, data = d, 
                family = poisson(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_pois)
```

Random intercept but GLM
```{r}
rand_intercept <- rnorm(N, mean=0, sd=0.3)

log_mean_contacts <- pop_intercept + log_rr_age[age_grp] + 
  log_rr_dategrp[date_grp] + log_rr_hhsizegrp[hhsize_grp] +
   + random_error + rand_intercept

contacts <- rtpois(N, lambda=exp(log_mean_contacts), a=0, b=60)
wt <- rnorm(N, 1, 0.2)

d <- data.frame(id = as.character(id),
                contacts = contacts, 
                age_grp = as.factor(age_grp), 
                hhsize_grp = as.factor(hhsize_grp),
                date_grp = as.factor(date_grp),
                wt = wt)

library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_pois <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp, data = d, 
                family = poisson(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_pois)
```


Random intercept, repeated measure but GLM
```{r}
nrepeat <- 7
rand_intercept <- rnorm(N, mean=0, sd=0.3)

log_mean_contacts <- pop_intercept + log_rr_age[age_grp] + 
  log_rr_dategrp[date_grp] + log_rr_hhsizegrp[hhsize_grp] +
   + random_error + rand_intercept

contacts <- as.vector(replicate(nrepeat, 
                      rtpois(N, lambda=exp(log_mean_contacts), a=0, b=60)))
wt <- rnorm(N, 1, 0.2)

d <- data.frame(id = rep(as.character(id), nrepeat),
                contacts = contacts, 
                age_grp = rep(as.factor(age_grp), nrepeat), 
                hhsize_grp = rep(as.factor(hhsize_grp), nrepeat),
                date_grp = rep(as.factor(date_grp), nrepeat),
                wt = rep(wt,nrepeat))

library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_pois <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp, data = d, 
                family = poisson(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_pois)
```

NegBin
```{r}
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_nb <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp, data = d, 
                family = negbinomial(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_nb)
```

Random intercept, repeated measure, mixed model
```{r}
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_pois_mix <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp + (1|id), data = d, 
                family = poisson(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_pois_mix)
```

Random intercept, repeated measure, mixed model
```{r}
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_nb_mix <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp + (1|id), data = d, 
                family = negbinomial(), iter=2000, cores = 4)
Sys.time() - tic
summary(fit_nb_mix)
plot(fit_nb_mix, variable = c("b_age_grp2"))
plot(conditional_effects(fit_nb_mix, effects = "age_grp:hhsize_grp"))
fixef(fit_nb_mix)
coef(fit_nb_mix)
pp_check(fit_nb_mix, ndraws=1e3)
loo(fit_pois_mix, fit_nb_mix, compare = TRUE)
loo_pois_mix <- loo(fit_pois_mix)
loo_nb_mix <- loo(fit_nb_mix)
loo_compare(loo_poix_mix, loo_nb_mix)
```

```{r}
contacts <- as.vector(
  replicate(nrepeat, rtpois(N, 
                            lambda=exp(log_mean_contacts),
                            a = 0, b = 60)))
wt <- rnorm(N, 1, 0.2)

d <- data.frame(id = as.character(as.vector(replicate(nrepeat, 1:N))),
                contacts = contacts, 
                age_grp = as.factor(rep(age_grp, nrepeat)), 
                hhsize_grp = as.factor(rep(hhsize_grp, nrepeat)),
                date_grp = as.factor(rep(date_grp, nrepeat)),
                wt = rep(wt, nrepeat))

head(d$contacts)
hist(d$contacts, breaks=30)
summary(d$contacts)


library(brms)
tic <- Sys.time()
set_prior("normal(0, 10)", class = "Intercept")
set_prior("normal(0, 2)", class = "b")

fit_pois <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp + (1|id), data = d, 
                family = poisson(), iter = 1000, warmup = 500, cores = 4)
Sys.time() - tic

tic <- Sys.time()
fit_negbin <- brms::brm(contacts|weights(wt) ~ age_grp + date_grp + 
                  hhsize_grp + (1|id), data = d, 
                family = negbinomial(), iter = 1000, warmup = 500, cores = 4)
Sys.time() - tic
```

Analysis of variance
```{r}
oneway_age <- aov(contacts ~ age_grp, data = dat[dat$date == as.Date("2023-12-06"),])
summary(oneway_age)
oneway_region <- aov(contacts ~ region, data = dat[dat$date == as.Date("2023-12-06"),])
summary(oneway_region)
oneway_hhsize <- aov(contacts ~ hhsize_grp, data = dat[dat$date == as.Date("2023-12-06"),])
summary(oneway_hhsize)
oneway_occupation <- aov(contacts ~ occupation, data = dat[dat$date == as.Date("2023-12-06"),])
summary(oneway_occupation)

age <- aov(contacts ~ age_grp, data = dat[dat$date == as.Date("2023-12-06"),])
summary(age)
ar <- aov(contacts ~ age + hhsize_grp, data = dat[dat$date == as.Date("2023-12-06"),])
summary(ar)
arh <- aov(contacts ~ age + hhsize_grp + region, data = dat[dat$date == as.Date("2023-12-06"),])
summary(arh)
arho <- aov(contacts ~ age + hhsize_grp + region + occupation, data = dat[dat$date == as.Date("2023-12-06"),])
summary(arho)
```


```{r}
d <- dat[,c("id","dayofweek", "contacts")]
dwide <- pivot_wider(d, names_from = dayofweek, values_from = contacts)
cov_obs <- select(dwide, ends_with("day")) %>% 
  cov()
cov_obs
cov2cor(cov_obs)
```

```{r}
library(GGally)
ggpairs(select(dwide, ends_with("day")), lower = list(continuous = "smooth"))
```






### Negative binomial vs. poisson

Negative binomial regression performs better in the likelihood ratio test and Akaike Information Criteria (AIC). Estimates are similar but
the standard errors will be biased for the poisson model

```{r}
fm1 <- contacts ~ age_grp + sex + dayofweek + hhsize_grp + region
m1_pois <- glm(fm1, data = dat, weights = wt, family="poisson")
m1_negbin <- glm.nb(fm1, data = dat, weights = wt)

knitr::kable(exp(m1_negbin$coefficients))
AIC(m1_pois, m1_negbin)
anova(m1_pois, m1_negbin)
rms::vif(m1_negbin)
```

### Negative binomial regression with varying covariates

```{r, eval=F, echo=F}
library(MASS)
fm_list <- list()

fm_list[[1]] <- contacts ~ age_grp + sex + dayofweek + hhsize_grp + region + occupation
fm_list[[2]] <- contacts ~ age_grp + sex + dayofweek + hhsize_grp + region
fm_list[[3]] <- contacts ~ age_grp + sex + dayofweek + hhsize_grp
fm_list[[4]] <- contacts ~ age_grp + sex + dayofweek
fm_list[[5]] <- contacts ~ age_grp + sex
fm_list[[6]] <- contacts ~ age_grp

glm_res <- lapply(fm_list, function(x) glm.nb(x, data = dat, weights = wt))
# saveRDS(glm_res, paste0("outputs/glm_res_", tstamp(), ".rds"))
```


Summary table 1
```{r}
glm_res <- readRDS("C:outputs/glm_res_20240521.rds")
df <- data.frame(
  Model = c("age_grp + sex + dayofweek + hhsize_grp + region + occupation", 
  "age_grp + sex + dayofweek + hhsize_grp + region", 
  "age_grp + sex + dayofweek + hhsize_grp", 
  "age_grp + sex + dayofweek", 
  "age_grp + sex", 
  "age_grp"), 
AIC = sapply(glm_res, function(x) AIC(x)))
knitr::kable(df)
anova(glm_res[[1]], glm_res[[2]])
rms::vif(glm_res[[1]])
knitr::kable(exp(glm_res[[2]]$coefficients))
```

POLYMOD-like table
```{r}
## Estimates
m <- glm_res[[2]]
xhat <- exp(summary(m)$coefficients[,1])
lb <- exp(summary(m)$coefficients[,1]) - 
  qnorm(0.975) * summary(m)$coefficients[,2] 
ub <- exp(summary(m)$coefficients[,1]) + 
  qnorm(0.975) * summary(m)$coefficients[,2]

est_ <- sprintf("%.2f (%.2f - %.2f)", xhat, lb, ub)
# include reference values
est <- c(est_[1], "1.00", est_[2:10], "1.00", est_[11], "1.00", est_[12:17],
"1.00",est_[18:21], "1.00", est_[22:37])


numpart <- c(age_counts$n, sex_counts$n, dayofweek_counts$n, 
             household_counts$n, region_counts$n)
mean_contacts <- c(age_counts$mean_contacts, sex_counts$mean_contacts,
                   dayofweek_counts$mean_contacts, 
                   household_counts$mean_contacts,
                   region_counts$mean_contacts)
sd_contacts <- c(age_counts$sd_contacts, sex_counts$sd_contacts,
                 dayofweek_counts$sd_contacts, household_counts$sd_contacts, 
                 region_counts$sd_contacts)

contacts <- sprintf("%.2f (%.2f)", mean_contacts, sd_contacts)

covars <- c("0-4", "5-9", "10-14", "15-19", "20-29","30-39", "40-49",
               "50-59", "60-69", "70+",
               "Female", "Male",
               "Sun","Mon","Tue","Wed","Thu","Fri",
               "Sat", 
               "1", "2", "3", "4", "5+",
            "Seoul",
                 "Busan","Daegu","Incheon","Gwangju","Daejeon","Ulsan",
                 "Sejong",
                 "Gyeonggi","Gangwon",
                 "Chungbuk","Chungnam",
                 "Jeonbuk","Jeonnam",
                 "Gyeongbuk","Gyeongnam",
                 "Jeju")

df <- data.frame(category=c(rep("Age",10), 
                          rep("Sex",2),
                          rep("Day",7), rep("HH size",5),
                          rep("Region",17)),
                 covariate = covars, 
                 num = numpart,
                 marginals = contacts,
                 estimates = est[2:42]) # exclude intercept

names(df) <- c("Category","Covar", "No. part", 
               "Marginal mean (sd)",
               "Est")

knitr::kable(df)
m$theta
m$SE.theta
est <- m$theta
lb <- m$theta - qnorm(0.975)*m$SE.theta 
ub <- m$theta + qnorm(0.975)*m$SE.theta 
sprintf("Dispersion parameter, theta = %.2f (%.2f - %.2f)", est, lb, ub)
sprintf("alpha = %.2f (%.2f - %.2f)", 1/est, 1/ub, 1/lb)
```

### `dayofweek` analysis

```{r, eval=F}
dat <- subset(dat_full, date > as.Date("2024-01-01"))

wk <- c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")
glm_by_dayofweek <- vector("list", length(wk))

for (i in 1:length(wk)) {
  dat_day <- subset(dat, dayofweek == wk[i])
  glm_by_dayofweek[[i]] <- 
    MASS::glm.nb(contacts ~ age_grp + sex + hhsize_grp + region,
                             data = dat_day, weights = wt)
}

saveRDS(glm_by_dayofweek,  paste0("outputs/glm_dayofweek_2024_", tstamp(), ".rds"))

```

### Generalized linear model(GLM)

```{r, eval=FALSE}
tic <- Sys.time()
library(MASS)
m <- glm(contacts ~ age_grp + sex + dayofweek + hhsize_grp, 
           weights=wt, data=dat, 
           family=poisson(link = "log"))
Sys.time() - tic
saveRDS(m,
        paste0("outputs/glmm_pois_id_", tstamp(), ".rds"))


library(MASS)
m <- glmer(contacts ~ age_grp + sex + date_grp + hhsize_grp + (1|id), 
           weights=wt, data=dat_full, 
           family=poisson(link = "log"),
           control=glmerControl(calc.derivs = FALSE,
                                   optimizer="bobyqa",
                                   optCtrl = list(maxfun = 2e5)))
saveRDS(m,
        paste0("outputs/glmm_pois_dategrp_id_", tstamp(), ".rds"))
```

### Generalized linear mixed model(GLMM)

`id` random effect

```{r, eval=FALSE}
tic <- Sys.time()
library(MASS)
m <- glmer(contacts ~ age_grp + sex + dayofweek + hhsize_grp + (1|id), 
           weights=wt, data=dat, 
           family=poisson(link = "log"),
           control=glmerControl(calc.derivs = FALSE,
                                   optimizer="bobyqa",
                                   optCtrl = list(maxfun = 2e5)))
Sys.time() - tic
saveRDS(m,
        paste0("outputs/glmm_pois_id_", tstamp(), ".rds"))


library(MASS)
m <- glmer(contacts ~ age_grp + sex + date_grp + hhsize_grp + (1|id), 
           weights=wt, data=dat_full, 
           family=poisson(link = "log"),
           control=glmerControl(calc.derivs = FALSE,
                                   optimizer="bobyqa",
                                   optCtrl = list(maxfun = 2e5)))
saveRDS(m,
        paste0("outputs/glmm_pois_dategrp_id_", tstamp(), ".rds"))
```

`dayofweek` random effect

```{r, eval=FALSE}
library(MASS)
tic <- Sys.time()
m <- glmer(contacts ~ age_grp + sex + dayofweek + hhsize_grp, weights=wt, 
              data=dat,
              control=glmerControl(calc.derivs = FALSE,
                                   optimizer="bobyqa",
                                   optCtrl = list(maxfun = 2e5)))

# m <- glmer.nb(contacts ~ age_grp + sex + dayofweek + hhsize_grp +
#                 region + (1|dayofweek), weights=wt, data=dat,
#               control=glmerControl(calc.derivs = FALSE,
#                                    optimizer="bobyqa"))



m <- glmer.nb(contacts ~ age_grp + sex + hhsize_grp +
                region + (1|dayofweek), weights=wt, data=dat, 
              control=glmerControl(calc.derivs = FALSE,
                                   optimizer="bobyqa"))
Sys.time() - tic
saveRDS(m,
        paste0("outputs/glmm_dayofweek_fixed_missing_", tstamp(), ".rds"))
```


Estimates
```{r}
glm_res <- readRDS("outputs/glm_res_20240521.rds") # take the second, 2, standard model
glmm_id <- readRDS("outputs/glmm_id_20240522.rds")
# glmm_dayofweek <- readRDS("outputs/glmm_dayofweek_20240522.rds")

summary_est <- function(m) {
  xhat <- exp(summary(m)$coefficients[,1])
  lb <- exp(summary(m)$coefficients[,1]) - 
    qnorm(0.975) * summary(m)$coefficients[,2] 
  ub <- exp(summary(m)$coefficients[,1]) + 
    qnorm(0.975) * summary(m)$coefficients[,2]
    est_ <- sprintf("%.2f (%.2f - %.2f)", xhat, lb, ub)
  # include reference values
  est <- c(est_[1], "1.00", est_[2:10], "1.00", est_[11], "1.00", est_[12:17],
"1.00",est_[18:21], "1.00", est_[22:37])
  return(est)
}

est_glm <- summary_est(glm_res[[2]])
est_glmm_id <- summary_est(glmm_id)

numpart <- c(age_counts$n, sex_counts$n, dayofweek_counts$n, 
             household_counts$n, region_counts$n)
mean_contacts <- c(age_counts$mean_contacts, sex_counts$mean_contacts,
                   dayofweek_counts$mean_contacts, 
                   household_counts$mean_contacts,
                   region_counts$mean_contacts)
sd_contacts <- c(age_counts$sd_contacts, sex_counts$sd_contacts,
                 dayofweek_counts$sd_contacts, household_counts$sd_contacts, 
                 region_counts$sd_contacts)

contacts <- sprintf("%.2f (%.2f)", mean_contacts, sd_contacts)

covars <- c("0-4", "5-9", "10-14", "15-19", "20-29","30-39", "40-49",
               "50-59", "60-69", "70+",
               "Female", "Male",
               "Sun","Mon","Tue","Wed","Thu","Fri",
               "Sat", 
               "1", "2", "3", "4", "5+",
            "Seoul",
                 "Busan","Daegu","Incheon","Gwangju","Daejeon","Ulsan",
                 "Sejong",
                 "Gyeonggi","Gangwon",
                 "Chungbuk","Chungnam",
                 "Jeonbuk","Jeonnam",
                 "Gyeongbuk","Gyeongnam",
                 "Jeju")

df <- data.frame(category=c(rep("Age",10), 
                          rep("Sex",2),
                          rep("Day",7), rep("HH size",5),
                          rep("Region",17)),
                 covariate = covars, 
                 num = numpart,
                 marginals = contacts,
                 estimates = est_glm[2:42],
                 est_glmm_id = est_glmm_id[2:42]) # exclude intercept

names(df) <- c("Category","Covar", "No. part", 
               "Marginal mean (sd)",
               "Est", "Est_GLMM_id")

knitr::kable(df)
```



